---
title: "Regression Models Quiz 3"
author: "Alejandro"
date: "26 Januar 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

Consider the mtcrs data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.

```{r}
data("mtcars")
mdl <- lm(mpg ~ factor(cyl) + wt, mtcars)
summary(mdl)$coef
```

### 2.

Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as a possible confounding variable. Compare the effect of 8 versus 4 cylinders on mpg for the adjusted and unadjusted by weight models. Here, adjusted means including the weight variable as a term in the regression model and unadjusted means the model without weight included. What can be said about the effect comparing 8 and 4 cylinders after looking at models with and without weight included?.

```{r}
fit1 <- lm(mpg ~ factor(cyl), mtcars)
fit2 <- lm(mpg ~ factor(cyl) + wt, mtcars)

summary(fit1)$coef
summary(fit2)$coef

```

Answer: Holding weight constant, cylinder appears to have less of an impact on mpg than if weight is disregarded. 

### 3.

Consider the mtcars data set. Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.


```{r}
fit1 <- lm(mpg ~ factor(cyl) + wt, mtcars)
fit2 <- lm(mpg ~ factor(cyl) + wt + wt*factor(cyl), mtcars)
anova(fit1,fit2)
```

Answer: The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms may not be necessary.

### 4. 

Consider the data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight inlcuded in the model as:
lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
How is the wt coefficient interpretted?

Answer: 

### 5.

Consider the following data set
```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
```
Give the hat diagonal for the most influential point

```{r}
fit <- lm(y~x)
max(hatvalues(fit))
x[which.max(hatvalues(fit))]
```


Consider the following data set

```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
```
Give the slope dfbeta for the point with the highest hat value.

```{r}
fit <- lm(y~x)
hv <- hatvalues(fit)
max.hv <- which.max(hv)
dfb <- dfbetas(fit)
dfb[max.hv,2]
```


### 7.
Consider a regression relationship between Y and X with and without adjustment for a third variable Z. Which of the following is true about comparing the regression coefficient between Y and X with and without adjustment for Z.

Answer: It is possible for the coefficient to reverse sign after adjustment. For example, it can be strongly significant and positive before adjustment and strongly significant and negative after adjustment.