---
title: "Statistical Inference Course Project Part 2"
author: "Alejandro"
date: "14 Januar 2018"
output: pdf_document
---

```{r setoptions, echo=FALSE, warning = FALSE, message=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen=999, comment = NA)
```


## Overview
In this Peer-Graded-Assignment the central limit theorem (CLT) and the ToothGrowth data will be analyzed. For the first part we will rely in the simulation power of R to empirically derive the CLT. In the second part the ToothGrowth data in the R datasets package will be analyzed using techniques learned in class such as confidence intervals and/or hypothesis tests.

```{r loadPackage, echo=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
library(reshape2)

```

## Part 2 

In this second part we will do statistical inference on the ToothGrowth data set from R. First we will take a look at the data set.
```{r}

data("ToothGrowth")
str(ToothGrowth)

```

We see that there is a total of 60 observations with 3 variables: len, supp, and dose. Variables len and dose are numerical variables while supp gibe us information about what supplement was use during the test being a factor variable with 2 levels. Using the summary function in R we can analyze the numeric variables and see their spread.

```{r}
summary(ToothGrowth)
```

We can now see a boxplot for both dose to give us an better picture of the data. Furthermore we can see the effect that the dose has in the length. 

```{r, fig.height=2, fig.width=3}
aggregate(len ~ dose, ToothGrowth, mean)
g <- ggplot(data = ToothGrowth, aes(x = supp, y = len, fill = supp))
g + geom_boxplot()
```

We see that the dose has a clear effect in the len. From the boxplot we can see that the supp changes the mean and the spread of the data. By using the t.test function in R we can compute the effect that both supplements have in the tooth growth. 

```{r}
t.output <- t.test(len ~ supp , data = ToothGrowth, paired = FALSE, var.equal = FALSE)
t.output$p.value
```
With a p value of 0.06 we cannot reject the null hypothesis with a confidence of 95% that the different supplements have an effect in the tooth growth.

```{r}
sub1 <- ToothGrowth[ToothGrowth$dose == 0.5 | ToothGrowth$dose == 1,]
sub2 <- ToothGrowth[ToothGrowth$dose == 1 | ToothGrowth$dose == 2,]

res1 <- t.test(len~dose, data = sub1, paired = FALSE, var.equal = FALSE)
res2 <- t.test(len~dose, data = sub2, paired = FALSE, var.equal = FALSE)


```

By looking at the pvalues: `r res1$p.value` and `r res2$p.value`, we can reject the null hypothesis that the dose does not have an effect in the length of the tooth growth with a 95% confidence. 

### Conclussion
We can conclude that different doses lead to an increase in tooth growth, but we cannot say that different supplements lead to a difference in tooth growth.